|

### 2021-08-16 appendix/appendix.Rmd
appendix first quarter

---
title: |
       | Supplementary Materials for "Legislature Size and Welfare: Evidence from Brazil"
author:
  - Umberto Mignozzetti[^umberto]
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

@@

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors $\mathbb{x}$.

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place, where the decision leaves the mayor's hands and either immediately stops or transfers to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the councilors in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan concerns; a \textit{partisan} reversal mechanism, where the council is divided in support and opposition for the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan concerns with non-partisan rent extraction. In the main paper, we report only the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that partisan a mechanism with only politics (city councilors have parties, that generates political benefits and costs), and an intermediate mechanism (a more realist mechanism, including both extremes).

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals, and she has to use the policy instruments at her disposal. These instruments will vary depending on the mechanism that we will investigate, but they are two in nature: targeted goods for the city councilor's constituency and private goods for the city councilor. The targeted goods improve the councilor's reelection chances while the private goods can be understood as corruption, aimed at ``buying off'' the support of a recipient councilor. The use of the policy instruments, will generate a cost of govern, $C_G$, that will vary with the policy choices and the legislature size.

Finally, the mayor has to propose a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the utility from rents and benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of reelection $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office. The utility is multiplied by an indicator $\mathbb{I}$ that captures whether the mayor's proposal was accepted by the city council.\footnote{The indicator function is defined as
\[
I(condition) \ = \ \twopdo{1}{\text{Condition is satisfied.}}{0}.
\]
} The expected utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ \mathbb{I}(\text{Approval})(u(r) + B_M\pi(g))
\]

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R$ resources, and we assume the municipality cannot run debts. Let $C_G$ the costs of govern. The budget balance constraint requires the offers that the mayor makes to the city councilors to satisfy the following rule:

@@

The expected utility for the city councilors depend on the mechanism that will determine the reversal policy. We explore three different mechanisms. First, in the *non-partisan legislature* mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. Resources diminish by a factor of $\delta \in (0,1)$, and the selected councilor will have to make a proposal. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process than repeats, until one proposal is finally accepted.

In the second reversal mode, the *partisan legislature* mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor belongs to a party aligned with the mayor, we say that she belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, the city councilor belongs to the opposition $O \not\subset G$. The party affiliation play an essential role in the *partisan legislature* mechanism: if a city councilor that belongs to the mayor's coalition rejects the mayor's proposal, she loses the election. Whereas, if the city councilor belongs to the opposition, she loses the election by supporting mayor's proposal. Although these assumptions are strong, they capture the main underlying political dynamics. Partisanship means in most political interactions that if a city councilor aligns with the mayor, her electorate has preferences toward policies that the mayor would implement. Voting against the mayor's proposals would represent to defy the partisan brand.

In the last reversal mode, which is the reversal mechanism we use in the paper, we combine both partisanship with legislature bargaining rents. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the *non-partisan legislature* mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. Opposition members, on the other hand, receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as city councilors do not control the Executive, no public goods are provided ($p=0$). Moreover, in equilibrium, we assume that city councilors would prefer to receive rents instead of providing public goods.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, \bf{x})_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, we will have three different reversal policies. We will detail them below.
\end{enumerate}

This is a bargaining game, along the lines of @rubinstein1982perfect. For a political science model similar to this model, see @baron1989bargaining.

The solution for this game consists in finding a Sub-game Perfect Nash Equilibrium using backward induction. In the model with infinitely repeated proposals, we extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium is based on the assumption that if a politician accepts an offer at time $t+1$, she should accept the same offer at time $t$. This gives us an objective of finding the offer that will be accepted at any point in the game. Applying this reasoning, we can find the optimal offer at $t=0$, which represents no delay in policy implementation.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors that for the mayor, and the city councilors, indexed by the round of the implementation. The mayor's policy vector is comprised of a public goods provision $g^*$, rents $r^*$, $\mathbb{t}$ XXXX
\end{definition}

### The Mayor's Stage

Assume that we have worked out the solution using backward induction, and when there are $N$ councilors, we find that the cost of govern for the mayor is $C_G(N)$. This assumption will be dropped when we work each of the cases below.

We now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as possible for the public goods, and extract the remaining resources as political rents. Her objective is to maximize the following expected utility, subject to the municipal budget constraint.

@@

In equilibrium the budget binds, and the optimal public goods provision maximizes:

@@

\begin{proposition}
The provision of public goods increase with legislature size if the costs of govern decreases in larger legislatures.
\end{proposition}

\begin{proof}
In order to prove this result, we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a derivative on $N$ and use the implicit function theorem. Instead, we use monotone comparative statics [@milgrom1994monotone] to derive these results.

The mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1>N$, we have:

@@

And after substituting, we have:

\[
u(R - g - C_G(N+1)) + B_M\pi(g) - u(R - g' - C_G(N+1)) - B_M\pi(g')
\\
\geq 
\\
u(R - g - C_G(N)) + B_M\pi(g) - u(R - g' - C_G(N)) - B_M\pi(g')
\]

@@

This is true when the cost of govern with a council size of $N+1$ is lower than the cost of govern when the council size has $N$ councilors:

@@

Therefore, to satisfy increasing differences in the public goods provision and in the legislature size, the costs of govern has to decrease with legislature size.
\end{proof}

This provides us our first empirical hypothesis to test: *if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase*.

@@

### Non-Partisan Legislature

In the *non-partisan legislature* mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. Suppose that there were $k$ rejections, and we are at the $k+1$-th stage in the game. A given councilor will accept the proposer's offer if, and only if, accept the offer is better than wait until the next stage. If the offer is $x_i$, then:

@@

In the left-hand side, we place the offer. In the right-hand side, there are two components. The first is the amount the councilor $i$ gets when she is the proposer. It is equal to the budget in the next round minus the offers that she has to make to convince half of the councilors. The second part is the gains if she happens to reject but still receive an offer in the following round. Note that the offer she makes when she is the proposer is the same as the offer she wants to receive. This because the city councilors are exchangeable, and the solution is symmetric for all councilors receiving an offer (this means that we could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

@@

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the cost of govern is equal to:

@@

\begin{proposition}
In the *non-partisan legislature* mechanism, the costs of govern increase as $N$ increases.
\end{proposition}

@@

@@

Therefore, the costs always increase in $N$.

@@

\begin{corollary}
In the *non-partisan legislature*, public goods provision decreases as the legislature size increases.
\end{corollary}

This result is crucial to understand the role of the politics in our model. If the costs of govern increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In our case, we show that there is a *political* reason for increasing the public goods provision.

### Partisan Legislature

We now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a value of $-p$. Moreover, as in the case of the non-partisan legislature, the mayor can also distribute $x_i$ private goods for the legislator $i$. In this case, when the policy is implemented, utility of an aligned legislator is equals to $x_i + p$. When the legislator is not aligned with the mayor, her utility is equals to $x_i - p$. If the policy is not implemented, then we assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor will always support the mayor's policy, regardless of any private goods. This because as $p > 0$, the councilor always receive positive benefits from the policy. However, a politician not aligned with the mayor will require a compensation if her vote is needed. In this case, if the mayor needs the support of an opposition politician to pass her policy proposal, then she will need to compensate her for the policy costs. In this case:

@@

Optimality dictates that the mayor will offer $x_i = p$ to an opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of govern when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor will have to compensate:

@@

\begin{proposition}
In the *partisan legislature* mechanism, if the chance of electing a mayor-aligned politician is greater than one-half, then the costs of govern decrease as $N$ increases.
\end{proposition}

@@

\begin{corollary}
In the *partisan legislature*, if the chance of electing a mayor-aligned politician is greater than $1/2$, then public goods provision increase as the legislature size increases.
\end{corollary}

@@

### Hybrid Partisan and Non-partisan Legislatures

The hybrid mechanism combines both partisan and non-partisan motivations. In the hybrid mechanism, a city councilor aligned with the government favors the mayor's offer if:

@@

An opposition politician, on the other hand, favors the mayor's offer when:

@@

Inspecting these equations we can decompose these costs into costs in terms of rents and costs (or benefits) from political alignment. In this context, the cost of govern will be a weighted average between the expected payments for councilors that belong to the mayor's coalition, versus the expected payments for the councilors that belongs to the opposition. The chance of a opposition member be elected is equal to $\gamma$, and taking the weighted averages, the costs become:

@@

\begin{proposition}
In the *hybrid legislature*, if $\gamma \leq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right]$, then the costs of govern increase as the legislature size increases. Otherwise, the costs decrease when the legislature size increases.
\end{proposition}

@@

@@

\begin{corollary}
In the *hybrid legislature* mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

Also, note that the threshold $\gamma$ for the probability of electing a politician aligned with the government is decreasing in both the political outcome $p$ and the size of the legislature $N$. This means that more political benefits from policy and larger legislatures makes it easier to satisfy the electoral constraint.

@@

Taking stocks, the models provide two empirically testable hypotheses. Consider a legislature that previously had $N$ legislators, but now needs to increase its legislature by $N+1$. Then:

@@

@@

>>>>

---
title: |
       | Supplementary Materials for "Legislature Size and Welfare: Evidence from Brazil"
author:
  - Gabriel Cepaluni[^gabriel]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
date: \today
fontfamily: libertine
fontawesome: yes
fontsize: 11pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
always_allow_html: true
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    toc: true
    keep_tex: no
    template: template.latex
---

@@

Consider an strategic interaction between a mayor ($M$) and $N \geq 3$ city councilors, indexed in $i \in \{1, 2, \cdots, N \}$. The mayor is the head of the local executive branch, and has the prerogative of proposing a vector of policies, that is voted by the city councilors. Policy proposals are a combination of public goods provision $g$, rents for the mayor $r$, and a vector of transfer for city councilors \textbf{x}.

The city council votes the mayor's proposal and if they accept it, it is implemented. Otherwise, a reversal policy takes place, where the decision leaves the mayor's hands and either immediately stops or transfers to the council. The types of reversal policies are crucial for our argument, as it changes the relative strength of the councilors in their bargaining with the mayor. We investigate three types of reversal policies: a \textit{non-partisan} reversal mechanism, where city councilors have no partisan concerns; a \textit{partisan} reversal mechanism, where the council is divided in supporters and non-supporters of the mayor; and finally a \textit{hybrid} reversal mechanism, that combines partisan concerns with non-partisan rent extraction. In the main paper, we report only the hybrid model, as it is the more realistic scenario. Despite that, comparing the partisan and non-partisan reversal mechanisms illustrate clearly our argument that the partisan a mechanism is what drives our results.

To govern, the mayor has to convince at least half of the city councilors to support her policy proposals. Similar to selectorate models \citep{buenodemesquita2005logic}, we assume that the selectorate, i.e. the individuals that have a say in the mayor's policy proposals, are the $N$ city councilors. The winning coalition, which is the minimum number of members that have to be convinced to implement the policy, is equal to half of the selectorate $N/2$. City councilors are moved by policy concerns $p$ and transfers received from the mayor \textbf{x}. The use of transfers to affect the councilors' choices generate a cost of govern, $C_G$, that vary with the reversal mechanism and the legislature size.

Finally, after paying the costs of govern, the mayor proposes a level of public goods provision ($g$) and rents ($r$). Public goods provision help the mayor to get reelected. Rents are for direct mayor's consumption, and does not contribute to the mayor's electoral success. This makes the mayor's expected utility a sum of the gains from rents and the benefits from reelection. Both the utilities from rents ($u(r)$) and the probability of reelection $\pi(g)$ are concave functions, meaning that more rents or public goods increase the utility at a decreasing rate. The probability of reelection is multiplied by the benefit from holding office $B_M > 0$. This benefit captures the tangible and intangible gains that the mayor perceives from holding the public office, and we assume that they are high enough to rule out an equilibrium with only rents and transfers. If the mayor's policy proposal is approved, then the expected utility for the mayor is:

\[
\mathbb{E}U(r, g) \ = \ u(r) + B_M\pi(g)
\]

The policy choices of the mayor are subject to the municipal budget constraint. The municipality has $R > 0$ resources, and cannot run debts, meaning that the budget has to be balanced. We also assume that there are enough resources for the mayor to govern.

Let $C_G$ the costs of govern. The budget balance constraint requires the offers that the mayor makes to the city councilors satisfy the following inequality:

@@

The expected utility for the city councilors depend on the mechanism that will determine the reversal policy. As discussed, we explore three different mechanisms. First, in \textit{non-partisan} reversal mechanism, when the council rejects the mayor's proposal, a reversal stage starts with resources diminish by a factor of $\delta \in (0,1)$ and the selection of one councilor randomly. The randomly selected councilor becomes now the proposer. If her proposal is accepted, it is implemented. If it is rejected, the budget shrinks again by a factor $\delta$, and another city councilor is recognized to make the proposal. The process repeats, until one proposal is finally accepted.

In the second reversal mode, the \textit{partisan} reversal mechanism, we assume that each councilor has a party affiliation. Party affiliations are mutually exclusive (a councilor cannot belong to two parties at once). If a city councilor is affiliated to a party aligned with the mayor, we say that she belongs to the government coalition ($G \subset \{1, 2, \cdots, N \}$). Otherwise, the city councilor belongs to the opposition $O \not\subset G$. The mayor's policy generate a political value of $p > 0$ for the councilors, and whether it adds or subtracts from the councilor's utility depends of the party affiliation of the councilor. If the mayor's proposal is rejected, a city councilor aligned with the mayor, that would receive $p$ from the policy, now receives zero. On the other hand, a city councilor that belongs to the opposition would receive $-p$ from the mayor's policy, and now receives zero. Although these assumptions are strong, they capture the underlying political dynamics of partisanship: the political preferences in terms of public goods provision are shared by councilors aligned with the mayor. Voting against the mayor's proposals would represent to defy the partisan brand.

In the last reversal mechanism, the \textit{hybrid} reversal mechanism, we combine partisanship with legislature bargaining rents. Councilors in this mechanism receive a partisan utility from the mayor's policy, but if they reject the policy, a reversal policy similar to the non-partisan reversal mechanism takes place. In the mayor's proposal stage, city councilors that belong to the mayor's coalition receive a gain of $p > 0$ when the mayor's policy is implemented. Opposition members, on the other hand, receive $-p$ when the mayor's proposal is implemented. If the mayor's proposal is rejected, as city councilors do not control the Executive when the two branches are separated, no public goods are provided ($p=0$). Moreover, in equilibrium, we assume that city councilors would prefer to receive rents instead of providing public goods. This aligns with the empirical evidence in most developing economies.

@@

\begin{enumerate}
  \item The mayor learns how many government $|G|$ and opposition $|O|$ legislators were elected.
  \item The mayor proposes a policy vector $(r, g, x)_M$.
  \item The city council votes the proposal.
  \begin{itemize}
    \item If the council accepts, the policy is implemented and the game ends. 
    \item Otherwise, the reversal policy is implemented.
  \end{itemize}
  \item (Reversal Policy) Depends on which mechanism, we will have three different reversal policies. We will detail them below.
\end{enumerate}

This is a bargaining game, along the lines of \citet{rubinstein1982perfect}. For a political science models similar to this model, see \citet{baron1989bargaining} and, to some extent, \citet{buenodemesquita2005logic}.

The solution concept we use in this model is the Sub-game Perfect Nash Equilibrium. A Sub-game Perfect Nash Equilibrium requires that the strategies played are a Nash Equilibrium in each of the subgames of the game. The solution is found using backward induction. In the model with infinitely repeated proposals, we extend the equilibrium concept to require that the equilibrium is Stationary. A Stationary Sub-game Perfect Equilibrium is based on the assumption that, at each given point of the game, if a politician accepts an offer at time $k+1$, she should accept the same offer at time $k$. This gives us an objective of finding the offer that will be accepted at any point in the game. Applying this reasoning, we can find the optimal offer at $k=0$, which represents no delay in policy implementation.

\begin{definition}
The solution for the game consists in a pair of strategic policy vectors $(g^*, r^*, x^*)_k$, for $k \in \{0, 1, \cdots, \infty\}$, such that, for the mayor and the councilors, there are no profitable deviations from the equilibrium.
\end{definition}

Moreover, we assume that the players will play no dominated strategies, even when there is less than sufficient support for implement the policy.

### The Mayor's Decision Stage

Solving the game by backward induction would require us to start with the councilor's strategies. However, for simplicity, we will assume at this point that we solved the the game for the councilors, and found the equilibrium cost of govern $C_G(N)$ when there are $N$ city councilors. As we will show in a few paragraphs, this is without loss of generality as we will drop the assumption. We use it to simplify the exposition.

We now derive the optimal rents ($r$) and public goods provision ($g$) proposed by the mayor. The mayor benefits from public goods provision, as it increases the chances of her reelection. However, she prefers to contribute as minimum as necessary for the public goods, and extract the remaining resources as political rents. Her objective is to maximize the following expected utility, subject to the municipal budget constraint.

@@

In equilibrium the budget binds, and the public goods provision maximizes:

@@

\begin{proposition}
The provision of public goods increase with legislature size when the costs of govern decrease with legislature size.
\end{proposition}

\begin{proof}
To prove this result we need to show that the mayor's expected utility satisfy the increasing differences in $g$ and $N$. This would mean that when increasing the size of the council, the optimal solution $g^*$ would also increase. However, as $N$ increases discretely, we cannot take a the cross-partial derivative on $N$ or even use the implicit function theorem. Instead, we use monotone comparative statics to derive these results \citep{milgrom1994monotone}.

To start, note that $\mathbb{R} \times \mathbb{N}$ is a complete lattice. In this case, mayor's expected utility satisfy the increasing differences in $g$ and $N$ when, for $g > g'$ and $N+1 > N$, we have:

@@

After substituting, we have:

\[
u(R - g - C_G(N+1)) + B_M\pi(g) - u(R - g' - C_G(N+1)) - B_M\pi(g') \geq 
u(R - g - C_G(N)) + B_M\pi(g) - u(R - g' - C_G(N)) - B_M\pi(g')
\]

@@

And the function satisfies the increasing differences in $g$ and $N$ when the costs of govern an $N+1$-sized city council is lower than the costs of govern an $N$-sized city council:

@@

This provides us our first empirical hypothesis to test: if the cost of govern decreases when the legislature increases, then the public goods provision increase when the legislature size increase.

@@

### Non-Partisan Reversal Mechanism

In the \textit{non-partisan reversal} mechanism, we assume that when the council rejects the mayor's office, a reversal stage starts with the selection of one councilor randomly. To find the stationary sub-game perfect Nash equilibrium, suppose that there were $k-1$ rejections, and we are at the $k$-th stage in the game. A councilor will accept the proposer's offer if, and only if, to accept the offer is better than to wait until the next stage. If the offer is $x_i$, then:

@@

In the left-hand side, we place the offer. In the right-hand side, there are two components. The first is the amount that the councilor $i$ gets when she is the proposer. It is equal to the budget in the next round minus the offers that she has to make to convince half of the councilors. The second part is the gains if she happens to reject but still receive an offer in the following round. Note that the offer she makes when she is the proposer is the same as the offer she wants to receive. This because the city councilors are exchangeable, and the solution is symmetric for all councilors receiving an offer (this means that we could have dropped the $i$ in the solution). After some algebra, the offer $x_i$ has to be greater than or equal to:

@@

Proceeding backwards, at the mayor's proposal stage $k=0$ and the mayor is going to offer $\underline{x}(0, N) = \dfrac{2 \delta R}{2N + 1}$ to half of the councilors. In this context, the costs of govern are equal to:

@@

\begin{proposition}
In the \textit{non-partisan reversal} mechanism, the costs of govern increase as the size of the legislature increases.
\end{proposition}

@@

The difference in costs when there are $N+1$ city councilors versus when there are $N$ city councilors is equal to:

@@

Therefore, the costs always increase when the size of the legislature increases.

@@

\begin{corollary}
In the \textit{non-partisan reversal} mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

This result is crucial to understand the role of the politics in our model. If the costs of govern increase with legislature size, then any increase in public goods provision associated with legislature size has come from a different mechanism. In our case, we show that there is a \textit{political} reason for increasing the public goods provision.

### Partisan Reversal Mechanism

We now look into the other extreme, which is a fully politicized legislature. In this case, when the mayor implements the policy, it generates a value of $p > 0$. A legislator aligned with the mayor receives a benefit $p$ while an opposition legislator receives a cost of $-p$. Moreover, as in the case of the non-partisan reversal, the mayor can choose to distribute $x_i$ private goods for a legislator $i$. In this case, when the policy is implemented, the utility of an aligned legislator is equals to $x_i + p$. When the legislator is not aligned with the mayor, her utility is equals to $x_i - p$. If the policy is not implemented, then we assume a reversal policy of zero in all choice vectors $g = x_i = p = 0$.

A legislator aligned with the mayor will always support the mayor's policy, regardless of receiving or not private goods. This because as $p > 0$, the councilor always receive positive benefits from the policy.\footnote{In reality, if it was possible, the mayor could even impose a cost to the city councilors aligned with the government. Although this seem unrealistic, it is mostly the case when government-aligned legislators are recognized to draft legislation for the mayor's office.} However, a politician not aligned with the mayor will require a compensation if her vote is needed. In this case, if the mayor needs the support of an opposition politician to pass her policy proposal, then she will need to compensate her for the policy costs. In this case:

@@

Optimality dictates that the mayor will offer $x_i = p$ to the opposition legislator. Let $\gamma$ be the ex-ante probability of electing a politician aligned with the mayor. Then, the costs of govern when the legislature size is $N$ is equal to the expected number of politicians in the opposition that the mayor will have to compensate minus the amount needed to implement the policy, in this case, half of them:

@@

\begin{proposition}
In the \textit{partisan reversal} mechanism, if the chance of electing a mayor-aligned politician is greater than one-half, then the costs of govern decrease as the legislature size increases.
\end{proposition}

@@

\begin{corollary}
In the \textit{partisan reversal} mechanism, if the chance of electing a mayor-aligned politician is greater than $1/2$, then public goods provision increase as the legislature size increases.
\end{corollary}

@@

### Hybrid Partisan and Non-partisan Reversal Mechanism

The \textit{hybrid reversal} mechanism combines both partisan and non-partisan motivations. In the hybrid mechanism, a city councilor aligned with the government favors the mayor's offer if:

@@

An opposition politician, on the other hand, favors the mayor's offer if, and only if:

@@

Inspecting these equations, we can decompose these costs into costs in terms of rents and costs (or benefits) from political alignment. In this context, the cost of govern will be depend on the expected payments for councilors in the mayor's coalition, versus the expected payments for the councilors that belongs to the opposition. The chance of a opposition member be elected is equal to $\gamma$, and taking the weighted averages, these costs become:

@@

\begin{proposition}
In the \textit{hybrid reversal} mechanism, if $\gamma \leq \dfrac{1}{p}\left[ \dfrac{1}{(2N+1)(2N+3)} \right]$, then the costs of govern increase as the legislature size increases. Otherwise, the costs decrease when the legislature size increases.
\end{proposition}

@@

The difference in costs when there are $N+1$ city councilors versus the costs when there are $N$ councilors is equal to:

@@

\begin{corollary}
In the \textit{hybrid reversal} mechanism, public goods provision decreases as the legislature size increases.
\end{corollary}

Also, note that the threshold $\gamma$ for the probability of electing a politician aligned with the government is decreasing in both the political value of the mayors policy $p$ and the size of the legislature $N$. This means that higher political values and larger legislatures makes it easier to satisfy the electoral threshold.

@@

Taking stocks, the models provide two empirically testable hypotheses. Consider a city council comprised of $N$ legislators, that increase the council size to $N+1$. Then:

@@

In the paper, we show that the provision of education and healthcare improves in Brazil, and this results from the fact that the chances of electing a city councilor aligned with the mayor is of 91%.

@@

## Variable Sources and Descriptive Statistics
\label{appsources}

We use three information sources, either from Brazilian governmental agencies or available online. Table \ref{varsources} displays the primary sources and their respective URLs.

\begin{sidewaystable}[!htbp] \centering 
\caption{Data Sources} 
\label{varsources}
\scalebox{0.75}{
  \begin{tabular}{@{\extracolsep{5pt}}llll} 
  \\[-1.8ex]\hline\hline \\[-1.8ex] 
  Source Code & \multicolumn{1}{c}{Source Name} & \multicolumn{1}{c}{Description} & \multicolumn{1}{c}{URL} \\ 
  \hline \\[-1.8ex] 
  & & & \\
  DataSUS & Brazilian Health Ministry Data Service & Collects data on Health Care & \url{http://www.datasus.gov.br} \\
  & & & \\
  IBGE & Brazilian Institute of Geography and Statistics & Collects data Geography, Economics and Demography & \url{http://www.ibge.gov.br} \\ 
  & & & \\
  INEP & Data Service of Ministry of Education & Collects data on performance of Education & \url{http://www.inep.gov.br} \\ 
  & & & \\
  InterLegis & Senate Legislative Data Service & Collect Data on Legislative in Brazil & \url{http://www.interlegis.leg.br} \\
  & & & \\
  IPEA & Brazilian Institute of Applied Economics & Collects data on the Economy & \url{http://www.ipeadata.gov.br} \\ 
  & & & \\
  MDS & Social Security Ministry & Collect data on Social Coverage and Effectiveness of Social Programs & \url{http://www.mds.gov.br} \\ 
  & & & \\
  TSE & Superior Electoral Tribunal & Collects data on elections results & \url{http://www.tse.jus.br} \\
  & & & \\
  \hline\hline \\[-1.8ex] 
  \end{tabular} 
}
\end{sidewaystable}

Now let us define each variable.

### Outcomes Aggregated at the Municipal Level

\begin{itemize}
\item \textbf{Number of Seats 2000}: Number of councilors in the municipality by the 2000 elections (source: TSE).
\item \textbf{Population 2000}: Municipal population according to the 2000 Brazilian Census (source: IBGE).
\item \textbf{Per-Capita GDP Census 2000 (in millions)}: Municipal per-capita GDP computed by the IBGE Census (source: IPEA).
\item \textbf{Proportion of Poverty Census 2000}: Proportion of poverty in the municipality defined as people living on less than R\$ 70.00 a day (source: MDS).
\item \textbf{Number of Seats 2004}: Number of city-councilors in a municipality according to the Electoral Justice decision (source: TSE).
\item \textbf{Infant Mortality 2005--2008}: Infant mortality computed as the number of children born alive that died before reaching one year of age divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Post-Natal Mortality Rate 2005--2008}: Infant mortality computed as the number of children born alive that successfully lived 28 days, but died before reaching one year of age, divided by the number of children born alive, multiplied by 1,000 (source: DataSUS)
\item \textbf{Coverage Family Health Program 2005--2008}: Number of families covered by the Family Health Program (source: DataSUS).
\item \textbf{Prop. Born w. 7+ Pre-Natal 2005--2008}: Proportion of children born who went to seven or more prenatal care consultations (source: DataSUS)
\item \textbf{Enrollment Elementary School 2005--2008}: Number of children enrolled in elementary schools, averaged by classroom size (K--4) (source: INEP).
\item \textbf{Enrollment Middle School 2005--2008}: Number of children enrolled in middle schools, averaged by classroom size (5--9) (source: INEP)
\item \textbf{Quality of Elementary School Index 2005--2008}: IDEB score averaged by schools in the municipality. IDEB scores are composed by the student's grades in math and language, multiplied by an indicator of the distortion between the year the child is supposed to be and the year the child is. The estimators are cleaned to avoid influences of schools and classroom specific effects (source: INEP, years: 2005, 2007).
\item \textbf{Quality of Middle School Index 2005--2008}: IDEB score averaged by schools in the municipality. IDEB scores are composed by the student's grade in math and language, multiplied by a score capturing the distortion between the year the child is supposed to be in, and the year the child is studying. The estimators are cleaned to avoid the influence of schools and classrooms specific effects. (source: INEP, years: 2005, 2007).
\item \textbf{Mayoral Pre-Electoral Coalition Size 2004}: Number of elected councilors in the mayoral pre-electoral coalition (source: TSE)
\item \textbf{Number of Appointed Bureaucrats 2005--2008}: Number of bureaucrats of the direct administration that were appointed to jobs in the municipality (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Career Bureaucrats 2005--2008}: Number of career bureaucrats in the municipality. Career bureaucrats cannot be fired by politicians and earn considerably better pensions after retirement (source: IBGE, years: 2005, 2006, and 2008)
\item \textbf{Number of Councilor's Appointed Assistants 2005}: Number of assistants appointed for the councilor cabinet (source: InterLegis)
\item \textbf{Number of Females Elected 2004}: Number of females elected to city council (source: TSE)
\item \textbf{Number of Non-Whites Elected 2004}: Number of non-whites elected to city council. We collected this data for municipalities less than 10,000 inhabitants away from the cutoffs (source: our compilation based on the TSE candidate pictures). 
\item \textbf{Competition per Seat 2004}: Number of people running for city-councilor divided by the number of city councilors in the municipality (source: TSE)
\item \textbf{Proportion Approved Legislation 2005}: Number of approved legislation in 2005 divided by the number of proposed legislation (source: InterLegis)
\end{itemize}

### Legislation Dataset

We coded the legislation approved by the councilors in 64 municipalities that are 10 thousand inhabitants away from the council size thresholds. We classified the legislation in five categories:

\begin{itemize}
\item \textbf{Local Public Goods}: Legislation that provides a service targeted at the individual, group, or geographical level.
\item \textbf{Public Goods or General Legislation}: Legislation that provides a service for the entire municipality, or approves a law that does not discriminate among citizens.
\item \textbf{Oversight}: Legislation that requests information to the mayor's office or the bureaucracy about the provision of services.
\item \textbf{Others}: Legislation that is not classified as Local Public Goods, Public Goods, or Oversight. Usually honors or procedures.
\item \textbf{Education and Health Care}: Legislation about Education or Health Care provision.
\end{itemize}

### Online 2016 Former City Councilors Survey

We run an online survey asking 174 councilors whether helping with [...service...] gives [none, few, some, or many] votes. The services were:
